---
title: "Calculating variables of interest"
author: "Dr Simon Jackson"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

This vignette will describe how to use cleaned user log files to calculate a set of variables determined as being initially useful.

## When are variables calculated

As a step in the data analysis process, calculating variables comes after the log files of user data have been read and cleaned. This is directly integrated into the `read_all` function. However, this vignette will go into greater detail about the functions used in this stage. To learn more about the reading and cleaning steps, please read the vignettes on [cleaning](cleaning.html) and [reading](reading.html) log files.

## Example data

To demonstrate, we'll use a sample set of user log files.

```{r, message = F}
# Create path to all user data (this is specific to this document, you'll need to create your own)
baseLoc <- system.file(package = "adapter")
extPath <- file.path(baseLoc, "extdata")
path_to_user_data <- file.path(extPath, "eg_user_list/")

# Load adapter package and read all user's log files
library(adapter)
d <- read_all(path_to_user_data, calculate = FALSE)  # Turn variable calculation off so that we can do it separately
```

We will use our list of user data, `d`, in forthcoming analyses.

For later comparisons, here is the data without variables calculated:

```{r}
d
```

## Functions for calculating variables

The adapter package provides functions to calculate specific variables for a single user's data:

- `calc_anyevent()` to create a boolean variable in events and streams tibbles indicating whether an event is taking place.
- `calc_speed()` to calculate instantaneous speed in a driver's streams tibble from their velocity vectors.
- `calc_distance()` to calculate the distance travelled in the sim world since the last time log from position vectors.

The package also provides a wrapper function that will take an entire list of user log file data and run any or all of the specific functions:

- `calc_variables()`

This wrapper function is what you will use most often.

## Using `calc_variables()`

Pass a cleaned user list to `calc_variables()` to add the variables calculated by the specific variable functions:

```{r}
calc_variables(d)
```

For example, notice that the drivers now have a speed variable in their streams tibbles.

Additional arguments allow you to disable the calculation of specific variables. For example, `calc_variables(d, speed = FALSE)` will not calculate the speed variable.
