---
title: "Cleaning log files"
author: "Dr Simon Jackson"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

This vignette will describe how to clean users' log files using [functions provided in the adapter package](#functions-for-cleaning-log-files).

## Logs to clean

Cleaning requires user log files to have been read into R in the appropriate format. For details about how to read in log files so that they can be cleaned using adapter functions, please read the [Vignette on Reading log files](reading.html).

To demonstrate, the following will read in a list of user log files using the `read_all()` function:

```{r, message = F}
# Create path to all user data (this is specific to this document, you'll need to create your own)
baseLoc <- system.file(package = "adapter")
extPath <- file.path(baseLoc, "extdata")
path_to_user_data <- file.path(extPath, "eg_user_list/")

# Load adapter package and read all user's log files
library(adapter)
d <- read_all(path_to_user_data, clean = FALSE, calculate = FALSE)  # Turn cleaning off so that we can do it separately
```

We will use our list of user data, `d`, in forthcoming analyses.

For later comparisons, here is the uncleaned data for one user in this list:

```{r}
d[[1]]
```

## Functions for cleaning log files

The adapter package provides the following functions for cleaning log files:

- [`clean_laps()`](#cleaning-with-clean_laps)
- [`clean_events()`](#cleaning-with-clean_events)

We'll cover these below.

## Cleaning with clean_laps()

The purpose of `clean_laps()` is to:

- Remove data before the first lap commences
- Remove data after the final lap has ended
- Add a `lap` column to `events` and `streams` tibbles to ease lap-based analysis (e.g., via `dplyr::group_by(lap)`).

To use `clean_laps()` most efficiently, pass it a list of user data created via `read_all`. For example, using `d` from above:

```{r, message = F}
clean_d <- clean_laps(d)
```

Lets now compare the data for the first user:

```{r}
clean_d[[1]]
```

If you compare to the uncleaned version above, you'll notice that there are fewer rows in the `events` and `streams` tibbles. This is because data before lap 1 started and after lap 5 ended has been removed. You'll also notice that there is `lap` column in both of these tibbles.

As a side note, `clean_laps()` will work on a single driver user's list. However, to handle drone operators, it must have the complete user list. This is because the drone operator information needs to come from his/her driver teammate (where lap information is logged originally).

## Cleaning with clean_events()

The purpose of `clean_events()` is to add boolean (logical `TRUE`/`FALSE`) columns to the events and stream tibbles that capture when certain events are happening. The function will add a new boolean column for each type of event, which will be `TRUE` at times during which the driver in the team was within the boundaries of the trigger box collider defining the event.

To use `clean_events()` most efficiently, pass it a list of user data created via `read_all`. For example, using `d` from above:

```{r, message = F}
clean_d <- clean_events(d)
clean_d[[1]]
```

Notice the boolean columns on the right for each event.
