---
title: "Typical-use Example"
author: "Dr Simon Jackson"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Typical-use Example}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

This vignette walks through a typical case of using the adapter package and how to best leverage its functions and features.

## Instllation and loading

Make sure that adapter is installed. For installation instructions, please read the vignette, ["Getting started"](getting.started.html).

Once installed, load adapter to make use of it:

```{r}
library(adapter)
```

## Reading data

You will almost always begin by reading in the simulation log file data. This is best done by having the data stored in a specific directory structure. For more information about how to structure data or how to handle cases when the data is not structured this way, please read the vignette, ["Reading log files"](reading.html).

Assuming the log files are structured appropriately, read the data into R using `read_all()` as demonstrated below.

**WARNING: reading the data can take a long time.**

```{r, message = F}
# Define a path to the top-level log file directory.
# The code shown here is specific to this example document.
# You will need to use your own (e.g., path_to_user_data <- "data/")
baseLoc <- system.file(package = "adapter")
extPath <- file.path(baseLoc, "extdata")
path_to_user_data <- file.path(extPath, "eg_user_list/")

# Read all users' log files and save as an object `d`
d <- read_all(path_to_user_data)
```

By using `read_all()`, you have imported the data as a list object in a convenient way. Important checks of the data coding are made (e.g., that there is only one driver and one or no drone operators per team, and that there aren't any duplicate unique ids). Other useful preparation and variable calculations occur automatically. For example:

- Data is cleaned to include only relevant data. For example, log information occuring before the first lap started or after the fifth lap ended is removed.
- Users have their teammate's unique id (`uid`) stored for easy access, if they had a teammate.
- Additional columns are added to the data regarding `events` and `streams` including:
    - Variables for tracking what's happening such as a numeric `lap` column and boolean event columns like `fog` and `ice`.
    - Useful features variables calculated in the streams such as instantaneous speed or distance travelled.

For more detailed information about this, please read the vignette, ["Reading log files"](reading.html).

### Saving time by saving the data

Reading the data can take a long time, and it's probably not something you want to do everytime you start working. Instead, 

### Structure of the user list 

It is important to understand is the structure of the list of user data returned by `read_all()`. Let's take a quick look:

```{r}
class(d)
length(d)
```




