---
title: "Worked Examples"
author: "Dr Simon Jackson"
date: "`r Sys.Date()`"
output:
  rmarkdown::html_vignette:
    toc: true
vignette: >
  %\VignetteIndexEntry{Typical-use Example}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

This vignette walks through the key functions and features of the adapter package, and provides worked examples that demonstrate the use of adapter in a proper analysis setting.

# Installation and loading

Make sure that adapter is installed. For installation instructions, please read the vignette, ["Getting started"](getting.started.html).

Once installed, load adapter to make use of it:

```{r}
library(adapter)
```

# Reading data

You will almost always begin by reading in the simulation log file data. This is best done by having the data stored in a specific directory structure. For more information about how to structure data or how to handle cases when the data is not structured this way, please read the vignette, ["Reading log files"](reading.html).

Assuming the log files are structured appropriately, read the data into R using `read_all()` as demonstrated below.

**WARNING: reading the data can take a long time.**

```{r, eval = F}
# Define a path to the top-level log file directory.
# The code shown here is specific to this example document.
# You will need to use your own (e.g., path_to_user_data <- "data/...")
baseLoc <- system.file(package = "adapter")
extPath <- file.path(baseLoc, "extdata")
path_to_user_data <- file.path(extPath, "eg_user_list/")

# Read all users' log files and save as an object `d`
d <- read_all(path_to_user_data)
```

By using `read_all()`, you have imported the data as a list object in a convenient way. Important checks of the data coding are made (e.g., that there is only one driver and one or no drone operators per team, and that there aren't any duplicate unique ids). Other useful preparation and variable calculations occur automatically. For example:

- Data is cleaned to include only relevant data. For example, log information occuring before the first lap started or after the fifth lap ended is removed.
- Users have their teammate's unique id (`uid`) stored for easy access, if they had a teammate.
- Additional columns are added to the data regarding `events` and `streams` including:
    - Variables for tracking what's happening such as a numeric `lap` column and boolean event columns like `fog` and `ice`.
    - Useful features variables calculated in the streams such as instantaneous speed or distance travelled.

For more detailed information about this, please read the vignette, ["Reading log files"](reading.html).

## Saving time by saving data

Reading the data can take a long time, and it's probably not something you want to do everytime you start working. Instead, it's advisable that after reading the data once, you save it as a properly formatted R object that can be easily read back into R on future occassions. Here is a demonstration of how this can be done by saving the list of user data in a .RDS format into a directory, "data/saved/".

```{r, eval = F}
# Define the directory in which to save
dir_to_save <- "data/save/"

# Define the file name (must end in `.rds`)
file_name <- "saved_user_list.rds"

# If the saved object doesn't exist, then create the directory (if needed) and
# save the object
if(!file.exists(file.path(dir_to_save, file_name))) {
  # Create a data/saved directory (unless it already exists)
  dir.create(dir_to_save, recursive = TRUE, showWarnings = FALSE)
  
  # Save object as RDS
  saveRDS(d, file.path(dir_to_save, file_name))
}
```

Now you can reload this saved object each time you start working as follows:

```{r, eval = F}
d <- readRDS(file.path(dir_to_save, file_name))
```

It's good practice to bundle all of these components together at the top of your script so you can read and save or load data as appropriate each time. You can do this as follows:

```{r}
# Define directory path and file name
dir_to_save <- "data/save/"
file_name <- "saved_user_list.rds"

# If the saved object doesn't exist, read is all data, create the directory (if 
# needed), and save the object. Otherwise, read in the saved object.
if(!file.exists(file.path(dir_to_save, file_name))) {
  
  # ** Replace these lines with a path to YOUR data directory **
  baseLoc <- system.file(package = "adapter")
  extPath <- file.path(baseLoc, "extdata")
  path_to_user_data <- file.path(extPath, "eg_user_list/")
  
  # Read all users' log files and save as an object `d`
  d <- read_all(path_to_user_data)
  
  # Create a data/saved directory (unless it already exists)
  dir.create(dir_to_save, recursive = TRUE, showWarnings = FALSE)

  # Save object as RDS
  saveRDS(d, file.path(dir_to_save, file_name))
} else {
  d <- readRDS(file.path(dir_to_save, file_name))
}
```

With the above, you can rerun your code multiple times but will only have to wait for the full reading of data once.

## Structure of the user list 

It is important to understand the structure of the list of user data returned by `read_all()`.

**It is a list**

```{r}
is.list(d)
```

**Each element contains information about a single user/participant**

```{r}
cat("Data is available for", length(d), "participants:", names(d))
```

**Information for each participant is a list**

```{r}
sapply(d, is.list)
```

**Each participant list contains three data frames**

```{r}
user_1 <- d[[1]]
cat("Each user contains", length(user_1), "data frames:", names(user_1), "\n\n")
sapply(user_1, is.data.frame)
```

These three data frames each contain different information:

- `session` contains session-level information like the user id and time the game started
- `events` contains timestamped logs of discrete events, like collisions, or the start/end of events.
- `streams` contains timestamped logs of variables collected every frame like the accelerator value.

For more information, please read the vignette, ["Reading log files"](reading.html).

# Working with the data

adapter provides a number of functions to support working with the data. This section will demonstrate their uses.

## Use the tidyverse

The adapter functions follow the philosophy of, and are best used in conjunction with, packages in the [tidyverse](http://tidyverse.org/). From here, we will be using tidyverse functions, which can be loaded as follows:

```{r, message = F, warning=F}
# install.packages("tidyverse")  # Install if needed
library(tidyverse)
```

It is recommended that tidyverse be loaded in your scripts.

```{r, echo = F}
# Quietly Set theme
theme_set(theme_minimal())
```

## Getting drivers or drones

You'll often want to separate data belonging to participants who were drivers or, less commonly, who were drone operators. Two functions are provided to handle these tasks: `get_drivers()` and `get_drones()`. Both functions take a list of user data and return a list in the same structure, but only with participants of the particular role.

For example:

```{r}
drivers <- get_drivers(d)
drones  <- get_drones(d)

cat("There were", length(drivers), "drivers (", names(drivers),
    ") and", length(drones), "drone operator(s) (", names(drones), ")")
```

## Binding a particular data frame

It's often the case that you will want to get a single data frame that contains the information for all participants (e.g., `session` or `streams`). adapter provides the function `bind_tbl()` to support this. To use it, you must specify your list of user data and the name of which data frame you want to bind (without quotes). For example:

```{r}
bind_tbl(d, session)
```

## Worked examples

This section contains examples of working with the sim data to accomplish various tasks.

### Mapping routes

This example demonstrates how to plot the routes taken by each driver, overlaid with the intended route. This is useful for explaining how the laps worked, as well as for identifying deviations from the intended route.

Mapping the intended routes is made simple because adapter comes with a data object containing information about the `waypoints` for each lap:

```{r}
waypoints
```

Each row of this tibble contains the lap number, and x and z coordinates of a waypoint (intended turn) in the simulation. You can read more detailed information on the help page by running `?waypoints`. With the added knowledge that laps always started and ended at the same point (approximately x = 600, z = 200), the intended route can be plotted using `waypoints` as follows:

```{r, fig.width=7.5}
ggplot(waypoints, aes(x = x, y = z)) +
  # Draw route path between waypoints
  geom_path(colour = "grey", size = 3, lineend = "round", alpha = .7) +
  # Draw arrow indicating beginning of each lap
  annotate("segment", x = 600, xend = 600, y = 200, yend = 300, color = "black", alpha = .5,
           arrow = arrow(type = "closed", angle = 30, length = unit(.1, "inches"))) +
  # Split routes across facets
  facet_wrap(~ lap, nrow = 1) +
  # Ensure x and y scales are equally spaced
  coord_equal() +
  # Provide title information
  ggtitle("Drivers' planned routes on each city lap")
```

The following challenge is to extract driver's actual routes The information can be obtained using `bind_tbl()` in combination with `get_drivers()` and dplyr and tidyr functions (in tidyverse):

```{r, fig.width=7.5}
# Get drivers' position information at each frame
# - `position` is a 3-element vector of the x,y,z coordinates
driver_positions <- d %>% 
  get_drivers() %>% 
  bind_tbl(streams) %>% 
  select(uid, lap, time, position)

driver_positions

# Extract the x and z coordinates at each frame
driver_positions <- driver_positions %>% 
  mutate(x = map_dbl(position, ~ .[1]),
         z = map_dbl(position, ~ .[3]))

# Make sure to order by time, per driver, so path is plotted correctly
driver_positions <- driver_positions %>% 
  arrange(uid, time)
  
driver_positions  

# Plot these paths
ggplot(driver_positions, aes(x = x, y = z)) +
  geom_path(aes(color = uid), alpha = .4) +
  facet_wrap(~ lap, nrow = 1) +
  coord_equal() +
  ggtitle("Drivers' routes on each city lap")
```

Here's an example combining these plots (seeing all drivers as the same colour):

```{r, fig.width=7.5}
ggplot(waypoints, aes(x = x, y = z)) +
  # Draw waypoint-defined route
  geom_path(colour = "grey", size = 3, lineend = "round", alpha = .7) +
  annotate("segment", x = 600, xend = 600, y = 200, yend = 300, color = "black", alpha = .5,
           arrow = arrow(type = "closed", angle = 30, length = unit(.1, "inches"))) +
  # Overlay drivers' routes
  geom_path(data = driver_positions, aes(group = uid), alpha = .2, color = "dodgerblue") +
  # Plot-general aspects
  facet_wrap(~ lap, nrow = 1) +
  coord_equal() +
  ggtitle("Overlay of drivers' routes on each city lap",
           subtitle = "Grey line shows intended route\nAll laps start and end at x = 600, z = 200")
```

### Counting numbers of drivers and drones

This demonstrates a method for counting the number of drivers and drone operators:

```{r}
# Bind the session tibbles
sessions <- bind_tbl(d, session)

# Print numbers of drivers and drones
sessions %>% count(role)
```

### Counting the number of collisions per driver

Here's an example of using `bind_tbl()` in combination with `get_drivers()` and dplyr functions (in tidyverse) to count the number of collisions each driver had:

```{r}
# Get events for all drivers
driver_events <- d %>% 
  get_drivers() %>% 
  bind_tbl(events)

# Count number collisions per driver
driver_events %>% 
  group_by(uid) %>% 
  summarise(n_collisions = sum(event == "collision"))
```

### Calculating steering variability on black ice per lap, per driver

This example demonstrates a method for calculating variance in the steering wheel use during the black ice event, for each lap, for each driver.

```{r}
# Get streams tibble for all drivers
streams <- d %>% 
  get_drivers() %>% 
  bind_tbl(streams)

# Refine to variables of interest
streams <- streams %>%
  select(uid, lap, time, input_horizontal, ice)

# Filter to only keep rows when drivers were on black ice
streams <- streams %>%
  filter(ice)

streams

# Summarise to compute variance of brake during each black ice event (per lap)
steer_on_ice <- streams %>%
  group_by(uid, lap) %>%
  summarise(steer_var = var(input_horizontal))

steer_on_ice

# Plot the results
ggplot(steer_on_ice, aes(x = lap, y = steer_var)) +
  geom_path(aes(color = uid)) +
  geom_point(size = 4) +
  labs(x = "Lap number", y = "Steering Variance") +
  ggtitle("Driver's steering variance during the black ice event")
```



